---
gtitle: "Statistical Inference Course Project"
author: "Pulan Yu"
date: "2/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1 Simulation

In this section, the exponential distribution in R will be investigated and compared with the Central Limit Theorem. 

The exponential distribution will be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. lambda was set to be 0.2 for all of the simulations. 

The following shows a thousand simulation with R

```{r}
    lambda = 0.2
    n = 1000
    hist(rexp(n,lambda))
```

The distribution of averages of 40 exponentials can be ploted by 

```{r}
    mns <- NULL
    for (i in 1:1000){
      mns <- c(mns,mean(rexp(40,lambda)))
    }
    mn <- mean(mns)
    sd <- sd(mns)
    
    hist(mns,freq = FALSE)
    curve(dnorm(x,mean=mn,sd=sd),col="darkblue", lwd=2, add=TRUE)
    abline(v = mn, col="blue", lwd=3)
    text(mn+0.3,0.05,"mean")
    abline(v = mn-sd, col="red", lwd=3)
    text(mn-sd,0.5,"mean-sd")
    abline(v = mn+sd, col="red", lwd=3)
    text(mn+sd,0.5,"mean+sd")
```
```{r echo=FALSE}
    #library(ggplot2)
    #mns_data <- as.data.frame(mns)
    #p<-ggplot(data=mns_data, aes(x=mns)) + 
    #geom_histogram(color="black", fill="red",binwidth = 0.5)
    #p + geom_vline(size=3,color="blue",xintercept = mean(mns))
```

<b>The theoretical mean is 1/lambda which is 5</b>, and the <b>sample mean</b> is

```{r}
   mean(mns)
```
which is <b><font color='blue'>the thick blue bar in the above figure</font></b>.

The theoretical standard deviation is also 1/lambda which is 5, so <b>the theoretical variance is 25</b>. The sample variance and standard deviation are

```{r}
  var(mns)
  sd(mns)
```
respectively.

<b><font color='red'>The standard deviation is showed in two red lines with the left being mean-sd and right being mean+sd</font></b>.

The distribution is closely to a gaussion distribution with which can be seen from the figure since the histogram overlay with the gaussion distribution well.

## Part 2: Basic Inferential Data Analysis

ToothGrowth dataset has three columns: len, supp and dos. The following code show the information about the three columns.

```{r ToothGrowth}
  summary(ToothGrowth)
  boxplot(ToothGrowth)
```

Supp has two values: VC and OJ respectively. The dose has five levels from 0.5,1.0,to 2.0.

```{r}
  vc <- ToothGrowth[ToothGrowth$supp =='VC',]
  boxplot(vc$len~vc$dos,col='red',xlab = 'dos',ylab='len')
  oj <- ToothGrowth[ToothGrowth$supp =='OJ',]
  boxplot(oj$len~oj$dos,col='blue',add = TRUE)
  #hist(vc$len~vc$dos,col='red')
```