---
gtitle: "Statistical Inference Course Project"
author: "Pulan Yu"
date: "2/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1 Simulation

In this section, the exponential distribution in R will be investigated and compared with the Central Limit Theorem. 

The exponential distribution will be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. lambda was set to be 0.2 for all of the simulations. 

The following shows a thousand simulation with R

```{r}
    lambda = 0.2
    n = 1000
    hist(rexp(n,lambda))
```

The distribution of averages of 40 exponentials can be ploted by 

```{r}
    mns <- NULL
    for (i in 1:1000){
      mns <- c(mns,mean(rexp(40,lambda)))
    }
    mn <- mean(mns)
    sd <- sd(mns)
    
    hist(mns,freq = FALSE)
    curve(dnorm(x,mean=mn,sd=sd),col="darkblue", lwd=2, add=TRUE)
    abline(v = mn, col="blue", lwd=3)
    text(mn+0.3,0.05,"mean")
    abline(v = mn-sd, col="red", lwd=3)
    text(mn-sd,0.5,"mean-sd")
    abline(v = mn+sd, col="red", lwd=3)
    text(mn+sd,0.5,"mean+sd")
```
```{r echo=FALSE}
    #library(ggplot2)
    #mns_data <- as.data.frame(mns)
    #p<-ggplot(data=mns_data, aes(x=mns)) + 
    #geom_histogram(color="black", fill="red",binwidth = 0.5)
    #p + geom_vline(size=3,color="blue",xintercept = mean(mns))
```

<b>The theoretical mean is 1/lambda which is 5</b>, and the <b>sample mean</b> is

```{r}
   mean(mns)
```
which is <b><font color='blue'>the thick blue bar in the above figure</font></b>.

The theoretical standard deviation is also 1/lambda which is 5, so <b>the theoretical variance is 25</b>. The sample variance and standard deviation are

```{r}
  var(mns)
  sd(mns)
```
respectively.

<b><font color='red'>The standard deviation is showed in two red lines with the left being mean-sd and right being mean+sd</font></b>.

The distribution is closely to a gaussion distribution with which can be seen from the figure since the histogram overlay with the gaussion distribution well.

## Part 2: Basic Inferential Data Analysis

ToothGrowth dataset has three columns: len, supp and dos. The following code show the information about the three columns.

```{r ToothGrowth}
  summary(ToothGrowth)
  boxplot(ToothGrowth)
```

The supp is a factor with two levels: VC and OJ. The dose has three values: 0.5, 1.0 and 2.0. 

The following code will do an exploeratory analysis

```{R}
  
  vc <- ToothGrowth[ToothGrowth$supp =='VC',]
  vchalf <- vc[vc$dose==0.5,]
  vcone <- vc[vc$dose==1.0,]
  vctwo <- vc[vc$dose==2.0,]
  par(mfrow=c(2,3))
  hist(vchalf$len)
  hist(vcone$len)
  sd(vcone$len)
  hist(vctwo$len)
  oj <- ToothGrowth[ToothGrowth$supp =='OJ',]
  ojhalf <- oj[oj$dose==0.5,]
  ojone <- oj[oj$dose==1.0,]
  ojtwo <- oj[oj$dose==2.0,]
  hist(ojhalf$len)
  hist(ojone$len)
  hist(ojtwo$len)
  
  mean(vchalf$len)
  sd(vchalf$len)
  mean(vcone$len)
  sd(vcone$len)
  mean(vctwo$len)
  sd(vctwo$len)
  mean(ojhalf$len)
  sd(ojhalf$len)
  mean(ojone$len)
  sd(ojone$len)
  mean(ojtwo$len)
  sd(ojtwo$len)
```


Supp has two values: VC and OJ respectively. The dose has five levels from 0.5,1.0,to 2.0.

```{r cache=TRUE}
  par(mfrow=c(1,2))
  vc <- ToothGrowth[ToothGrowth$supp =='VC',]
  boxplot(vc$len~vc$dos,col=(c("red","gold","darkgreen")),xlab = 'dos',ylab='len',main="VC")
  oj <- ToothGrowth[ToothGrowth$supp =='OJ',]
  boxplot(oj$len~oj$dos,col=(c("red","gold","darkgreen")),xlab='dos',ylab='len',main='OJ')
 
  
```

Since there are two variables: supp and dose, firstly, the effect of dose is investigated. 

1. the 0.5 level

```{r}
  vchalf <- vc[vc$dose==0.5,]
  ojhalf <- oj[oj$dose==0.5,]
  t.test(vchalf$len,ojhalf$len,var.equal=TRUE)
```

2. the 1.0 level

```{r}
  vcone <- vc[vc$dose==1.0,]
  ojone <- oj[oj$dose==1.0,]
  t.test(vcone$len,ojone$len,var.equal=TRUE)
```

3. the 2.0 level

```{r}
  vctwo <- vc[vc$dose==2.0,]
  ojtwo <- oj[oj$dose==2.0,]
  t.test(vctwo$len,ojtwo$len,var.equal=TRUE)
```

Then the effect of dose is investigated.

For VC supply

```{r}
  vchalf <- vc[vc$dose==0.5,]
  vcone <- vc[vc$dose==1.0,]
  vctwo <- vc[vc$dose==2.0,]
  t.test(vchalf$len,vcone$len,var.equal=TRUE)
  t.test(vchalf$len,vctwo$len,var.equal=TRUE)
  t.test(vctwo$len,vcone$len,var.equal=TRUE)
```

For OJ supply

```{r}
  ojhalf <- oj[oj$dose==0.5,]
  ojone <- oj[oj$dose==1.0,]
  ojtwo <- oj[oj$dose==2.0,]
  t.test(ojhalf$len,ojone$len,var.equal=TRUE)
  t.test(ojhalf$len,ojtwo$len,var.equal=TRUE)
  t.test(ojtwo$len,ojone$len,var.equal=TRUE)
```

<h2>Conclusions:</h2>
For both OJ and VC supply, the level of difference has effect on the tooth length since the p-value of all t-test are less then 0.05 at 95% confidence level. The null hypothesis is rejected, the level has effect on the length and higher does is better for tooth growth.

For the same level, at low dose level 0.5 and 1.0, the supply has effect on the tooth length and OJ is better for tooth growth. Since the p-value of 0.5 and 1.0 level are 0.005304 and 0.0007807 respectively, so the null hypothesis is rejected and supply has impact on the length. For 2.0 level, the p-value is 0.9637, the null hypothese is failed to be rejected at 95% confidence level. It suggests that at high dose, the supply may not have statistical significant impact on tooth length. 



