---
gtitle: "Statistical Inference Course Project"
author: "Pulan Yu"
date: "2/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1 Simulation

In this section, the exponential distribution in R will be investigated and compared with the Central Limit Theorem. 

The exponential distribution will be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. lambda was set to be 0.2 for all of the simulations. 

The following shows a thousand simulation with R

```{r}
    lambda = 0.2
    n = 1000
    hist(rexp(n,lambda))
```

The distribution of averages of 40 exponentials can be ploted by 

```{r}
    mns <- NULL
    for (i in 1:1000){
      mns <- c(mns,mean(rexp(40,lambda)))
    }
    hist(mns)
    
    library(ggplot2)
    mns_data <- as.data.frame(mns)
    p<-ggplot(data=mns_data, aes(x=mns)) + 
    geom_histogram(color="black", fill="red",binwidth = 0.5)
    p
    mean(mns)
```